---
title: "In-class Exercise 2"
author: "Foo Jia Yi Samantha"
date-modified: "August 26, 2024"
date: "August 26, 2024"
execute: 
  eval: true
  echo: true
  freeze: true
---

```{=html}
<button>In-class Exercise</button> <button>R</button> <button>sf</button>
```
# 1. Importing Libraries into R

In this exercise, we will use the following R packages:

-   **sf**, a relatively new R package specially designed to import, manage and process vector-based geospatial data in R.
-   **tidyverse** for tidying data (<https://tidyr.tidyverse.org/>)

```{r}
pacman::p_load(sf, tidyverse)
```

# 2. Importing Data Sets into R

We will first import the three geospatial data sets into R using [*st_read()*](https://r-spatial.github.io/sf/reference/st_read.html) of the **sf** package.

## 2.1 I**mporting** MP14_SUBZONE_WEB_PL (Web) **.shp Data**

```{r}
# Import shapefile
mpsz14_shp <- st_read(dsn = "data", layer = "MP14_SUBZONE_WEB_PL")
```

There are **3 ways** to explore the contents of a simple feature data frame like **`mpsz`**!

### 1) Using *st_geometry()*

The sf data.frame contains a **geometry** column, that is a list of class `sfc`. We can retrieve the geometry list-column in this case by `mpsz$geom` or `mpsz[[1]]`, but the more general way uses **`st_geometry()`**.

```{r}
# Retrieve geometry column 
st_geometry(mpsz14_shp)
```

> 🔎 **Observations:** This `MP14_SUBZONE_WEB_PL` file consists of **323 features,** consisting of multipolygon features.

### 2) Using *glimpse()*

We use **`glimpse()`** from the **dplyr** package to understand the data type of each fields.

E.g. `FMEL-UPD_D` field is in **date** data type and `X_ADDR`, `Y_ADDR`, `SHAPE_L` and `SHAPE_AREA` fields are in **double-precision values**.

```{r}
# Get data types 
glimpse(mpsz14_shp)
```

> 🔎 **Observations:** This `MP14_SUBZONE_WEB_PL` file consists of **323 rows** and **16 columns** with datatypes ranging from integers, characters, date and doubles.

### 3) Using *head()*

We use **`head()`** from the **base R** package to get the full information of the feature object **`mpsz`**. The `n` value indicates the no. of rows. I will use thiis method for this in-class exercise.

```{r}
# Display top 5 rows of the feature object 
head(mpsz14_shp, n=5)  
```

> 🔎 **Observations:** We can notice that the `MP14_SUBZONE_WEB_PL` file consists of **5 features** and **15 fields,** with the top 5 rows per column.

## 2.2 I**mporting** MP14_SUBZONE_WEB_PL (Web) **.kml Data**

We use the below code chunk to export **`mpsz14_shp`** sf data.frame into **kml** file which saves the file into our data folder.

```{r}
# Convert .shp file into .kml
st_write(mpsz14_shp, 
         "data/MP14_SUBZONE_WEB_PL.kml",
         delete_dsn = TRUE)

# Import KML file
mpsz14_kml = st_read("data/MP14_SUBZONE_WEB_PL.kml")

# Display top 5 rows of the feature object 
head(mpsz14_kml, n=5)  
```

> 💡 **Note:** **`delete_dsn = TRUE`** will help delete the original data before rendering it

## 2.3 Importing MP19_SUBZONE_WEB_PL (No Sea) .shp Data

```{r}
# Import shapefile
mpsz19_shp <- st_read(dsn = "data", layer = "MPSZ-2019")
```

> 🔎 **Observations:** We can notice that the data file consists of **332 features** and **6 fields,** and follows the WGS64 coordinate system. Here we can notice it uses the \`ESRI Shapefile' driver.

## 2.4 Importing MP19_SUBZONE_WEB_PL (No Sea) .kml Data

```{r}
# Convert .shp file into .kml
st_write(mpsz19_shp, 
         "data/MP19_SUBZONE_WEB_PL.kml",
         delete_dsn = TRUE)

# Import KML file
mpsz19_kml = st_read("data/MP19_SUBZONE_WEB_PL.kml")

# Display top 5 rows of the feature object 
head(mpsz19_kml, n=5)  
```

> 🔎 **Observations:** We can notice that the `data`file also consists of **332 features** and **6 fields,** and follows the WGS64 coordinate system, but it uses the **kml driver** accordingly.

# 3. Handling Coordinate Systems

Recall that the **geographic coordinate system (e.g., WGS84)** is not appropriate for analyses that involve distance/area. Hence, it's common for us to **transform the original data** to a **projected coordinate system**.

Let's take a look at the `mpsz19_shp` simple feature data frame. It shows that it is in the WGS84 coordinate system, i.e., geographic coordinate system.

```{r}
st_geometry(mpsz19_shp)
```

Now, we'll transform `mpsz19_shp`'s coordinate system from geographic (WGS84) to projected (SVY21).

```{r}
mpsz19_shp <- st_transform(mpsz19_shp, crs = 3414)

# Display top 5 rows of the feature object 
head(mpsz19_shp, n=5)  
```

> 🔎 **Observations:** Notice that the `mpsz19_shp` object has been updated to the **SVY21** **projected CRS**

# 4. Working with Population Data

For aspatial datasets like *`respopagsex2011to2020.csv`,* we will import into Rstudio using `read_csv()` function of **readr** package.

```{r}
popdata <- read_csv("data/respopagesextod2023.csv")
```

> 🔎 **Observations:** The *`respopagsex2023.csv`* data contains **100928 rows** and **7 columns**

Aggregate the data to prepare a data.frame showing population by Planning Area and Planning Subzone.

```{r}
popdata2023 <- popdata %>%
  group_by(PA, SZ, AG) %>%
  summarise(`POP` = sum(`Pop`)) %>%
  ungroup() %>%
  pivot_wider(names_from=AG,
              values_from=POP)

colnames(popdata2023)
```

```{r}
popdata2023 <- popdata2023%>%
  mutate_at(.vars = vars(PA, SZ),
            .funs = list(toupper))
```

> **💡 Note:** we use **`muttate()`** to convert all values to a common **UPPERCASE** before we do a left-join since R requires data of similar datatype when matching the data.

```{r}
# Check column names
colnames(mpsz19_shp)
colnames(popdata2023)

```

```{r}
# Join 'popdata2023' with 'mpsz19_shp'
mpsz_pop2023 <- left_join(mpsz19_shp, popdata2023,
                          by = c("SUBZONE_N" = "SZ"))
```

> **💡 Note:** The merge is done using a `left_join`, meaning that all rows from `mpsz19_shp` will be kept, even if there is no corresponding row in `popdata2023`.
>
> 🔎 **Observations**: The resulting data frame, `mpsz_pop2023`, will contain all the columns from `mpsz19_shp` plus any additional columns from `popdata2023` that match the `SUBZONE_N` column.

```{r}
# Join 'mpsz19_shp' with 'popdata2023'
pop2023_mpsz <- left_join(popdata2023, mpsz19_shp,
                          by = c("SZ" = "SUBZONE_N"))
```

> 🔎 **Observations**: The resulting data frame, `pop2023_mpsz`, will contain all the columns from `popdata2023` plus any additional columns from `mpsz19_shp` that match the `sz` column.
